---
title: "Introduction to Deep Learning"
subtitle: "Classification -- Single-label, Multi-class"
author: "Rick Scavetta"
output:
  html_document:
  fig_caption: true
  toc: true
  toc_float:
    collapsed: false
    smooth_scroll: false
  toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, eval = TRUE)

# Initialize packages
library(keras)
library(tidyverse)
```

# Session 1 {.tabset .tabset-fade .tabset-pills}

## Intro

### Learning Goals

Developing deep learning to two core questions in supervised learning: Classification and Regression. 

The UCI Abalone data-set is a small and easy starting point since it can be used for predicting age as either a categorical or continuous variable, leading to the 

### Outline

- What is a tensor and why use it?
- What is keras and what is its relationship to TensorFlow?
- What is the deep in deep learning? ANNs and densely-connected networks.
- The math of deep learning: Basics of matrix algebra, gradient descent, backpropagarion, chain rule.
- The four stages of Deep learning.
- Parameters and hyper-parameter.
- Functions distinguishing classification and regression: loss and optimizer functions.

### Functions in this session:

Basic `keras` functions:

| Function                   | Description                                       |
|:---------------------------|:--------------------------------------------------| 
| [`keras_model_sequential()`](https://www.rdocumentation.org/packages/keras/versions/2.2.0/topics/keras_model_sequential) | Keras Model composed of a linear stack of layers. |
| `layer_dense()`	           | Add a densely-connected NN layer to an output.    |
| `compile()`                | Configure a Keras model for training.             |
| `fit()`                    | Train a Keras model.                              |


In our last case study, Boston, we'll perform a regression to predict a continouous response variable from 13 predictor variables. To accommodate for this different analytical problem, we'll use:

- A new normalization for the input data, z scores,
- A new loss function, `mse`,
- A new metric, `mae`, and
- No final activation function (i.e. scalar).

And since we have a really small data set we'll have:

- A very simple network architecture, and
- K-fold crossvalidation.

## Install tensorflow 

It's only necessary to run this once. 

```{r install, eval = F}
# for GPU
# install_keras(tensorflow = "gpu")

# or CPU:
# install_keras() # for cpu
```

# Part 1: Data Preparation

## Obtain data

```{r}

# Import data
c(c(train_data, train_labels), c(test_data, test_labels)) %<-% dataset_boston_housing()

source("Boston_Z.R")
```


## Prepare the labels:

The target, response variable:

```{r strTargets}
str(train_labels)
str(test_labels)
```

```{r}

# train_labels %>% 
#   data.frame(x = .) %>% 
#   ggplot(aes(x)) +
#   geom_histogram(binwidth = 1)
# 
# hist(train_labels, breaks =  50, xlim = c(0,50))

# range(train_labels)
# range(test_labels)
# 
# sum(is.na(train_labels))
# sum(is.na(test_labels))

train_labels <- as.integer(cut(train_labels, seq(0,50,10))) - 1 
test_labels <- as.integer(cut(test_labels, seq(0,50,10))) - 1

table(train_labels)
table(test_labels)

```


```{r}
train_labels_vec <- to_categorical(train_labels)
test_labels_vec <- to_categorical(test_labels)
```

# Part 2: Define Network

## Define the network

Here we specify the final activation function. We're going to use the sigmoid activation function, which will return a single value. That matches the format of our labels.

```{r architecture}
network <- keras_model_sequential() %>% 
  layer_dense(units = 2^6, activation = "relu", input_shape = c(13)) %>% 
  layer_dense(units = 2^6, activation = "relu") %>% 
  layer_dense(units = 5, activation = "softmax")
```

## View a summary of the network

```{r summary}
summary(network)
```

## Compile

```{r compile}
network %>% compile(
  optimizer = "rmsprop",
  loss = "categorical_crossentropy",
  metrics = c("accuracy")
)
```

# Part 3: Validate our approach

Let's set apart 1,000 samples in our training data to use as a validation set:

```{r}
index <- 1:101

val_data <- train_data[index,]
train_data <- train_data[-index,]

val_labels_vec <- train_labels_vec[index,]
train_labels_vec = train_labels_vec[-index,]
```

Now let's train our network for 20 epochs:

```{r echo=TRUE, results = "hide", warning = FALSE}
history <- network %>% fit(
  train_data,
  train_labels_vec,
  epochs = 100,
  batch_size = 2^4,
  validation_data = list(val_data, val_labels_vec)
)
```

Let's display its loss and accuracy curves:

```{r}
plot(history)
```

The network begins to overfit after nine epochs. Let's train a new network from scratch for nine epochs and then evaluate it on the test set.

```{r, echo=TRUE, results='hide'}
network <- keras_model_sequential() %>% 
  layer_dense(units = 2^6, activation = "relu", input_shape = c(13)) %>% 
  layer_dense(units = 2^6, activation = "relu") %>% 
  layer_dense(units = 5, activation = "softmax")
  
network %>% compile(
  optimizer = "rmsprop",
  loss = "categorical_crossentropy",
  metrics = c("accuracy")
)

history <- network %>% fit(
  train_data,
  train_labels_vec,
  epochs = 25,
  batch_size = 2^4,
  validation_data = list(val_data, val_labels_vec)
)
```

# Part 5: Check output

Let's return to our original model using the vectorized data:

## Metrics

```{r metrics}
metrics <- network %>% evaluate(test_data, test_labels_vec)
```

```{r}
metrics
metrics$acc
# Error rate: incorrect calling
1 - metrics$acc
```

## Predictions

```{r predictions}
network %>% predict_classes(test_data[1:10,])
```

```{r allPredictions}
predictions <- network %>% predict_classes(test_data)
actual <- unlist(test_labels)
totalmisses <- sum(predictions != actual)
```

# Confusion Matrix

```{r confusion, echo = F}
suppressPackageStartupMessages(library(tidyverse))
# library(dplyr)
data.frame(target = actual,
           prediction = predictions) %>% 
  filter(target != prediction) %>% 
  group_by(target, prediction) %>%
  count() %>%
  ungroup() %>%
  mutate(perc = n/nrow(.)*100) %>% 
  filter(n > 1) %>% 
  ggplot(aes(target, prediction, size = n)) +
  geom_point(shape = 15, col = "#9F92C6") +
  scale_x_continuous("Actual Target", breaks = 0:4) +
  scale_y_continuous("Prediction", breaks = 0:4) +
  scale_size_area(breaks = c(2,5,10,15), max_size = 5) +
  coord_fixed() +
  ggtitle(paste(totalmisses, "mismatches")) +
  theme_classic() +
  theme(rect = element_blank(),
        axis.line = element_blank(),
        axis.text = element_text(colour = "black"))

```

# Session Info

```{r}
sessionInfo()
```

