---
title: "Introduction to Deep Learning"
subtitle: "Regression"
author: "Rick Scavetta"
output:
  html_document:
  fig_caption: true
  toc: true
  toc_float:
    collapsed: false
    smooth_scroll: false
  toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, eval = TRUE)

# Initialize packages
library(keras)
library(tidyverse)
```

# Session 1 {.tabset .tabset-fade .tabset-pills}

## Intro

### Learning Goals

Developing deep learning to two core questions in supervised learning: Classification and Regression. 

The UCI Abalone data-set is a small and easy starting point since it can be used for predicting age as either a categorical or continuous variable, leading to the 

### Outline

- What is a tensor and why use it?
- What is keras and what is its relationship to TensorFlow?
- What is the deep in deep learning? ANNs and densely-connected networks.
- The math of deep learning: Basics of matrix algebra, gradient descent, backpropagarion, chain rule.
- The four stages of Deep learning.
- Parameters and hyper-parameter.
- Functions distinguishing classification and regression: loss and optimizer functions.

### Functions in this session:

Basic `keras` functions:

| Function                   | Description                                       |
|:---------------------------|:--------------------------------------------------| 
| [`keras_model_sequential()`](https://www.rdocumentation.org/packages/keras/versions/2.2.0/topics/keras_model_sequential) | Keras Model composed of a linear stack of layers. |
| `layer_dense()`	           | Add a densely-connected NN layer to an output.    |
| `compile()`                | Configure a Keras model for training.             |
| `fit()`                    | Train a Keras model.                              |


In our last case study, Boston, we'll perform a regression to predict a continouous response variable from 13 predictor variables. To accommodate for this different analytical problem, we'll use:

- A new normalization for the input data, z scores,
- A new loss function, `mse`,
- A new metric, `mae`, and
- No final activation function (i.e. scalar).

And since we have a really small data set we'll have:

- A very simple network architecture, and
- K-fold crossvalidation.

## Install tensorflow 

It's only necessary to run this once. 

```{r install, eval = F}
# for GPU
# install_keras(tensorflow = "gpu")

# or CPU:
# install_keras() # for cpu
```

# Part 1: Data Preparation

## Obtain data

```{r data, warning = FALSE}
c(c(train_data, train_targets), c(test_data, test_targets)) %<-% dataset_boston_housing()
```

## Examine data:

Our predictor variables:

```{r strDataPre}
str(train_data)
str(test_data)
```

The target, response variable:

```{r strTargets}
str(train_targets)
```

## Prepare the data:

Convert z-scores:

$$z_i=\frac{x_i-\bar{x}}{s}$$
```{r zScores, cache = T}
# parameters for Scaling:
mean <- colMeans(train_data) # mean of each column
std <- apply(train_data, 2, sd) # stdev of each column

# Calculate feature-wise (within-variable) z-scores: (x - mean)/std
train_data <- scale(train_data, center = mean, scale = std)
test_data <- scale(test_data, center = mean, scale = std)

```

# Part 2: Define Network

## Define the network as a function

In contrast to our previous case studies, we're going to call the same model multiple times. So we'll create a function with no arguments that we can call to create our model when ever we want to use it for training. 

Here, I've hardcoded the number of features for this dataset (`13`). To generalize, we could just use `dim(train_data)[2]` to get the number of dimensions from the training set.  

```{r defModel}
build_model <- function() {
  network <- keras_model_sequential() %>% 
    layer_dense(units = 64, activation = "relu", input_shape = 13) %>% 
    layer_dense(units = 64, activation = "relu") %>% 
    layer_dense(units = 1) 
    
  network %>% compile(
    optimizer = "rmsprop", 
    loss = "mse", 
    metrics = c("mae")
  )
}
```

Note two new functions here, the mean squared error:

$$\operatorname{MSE} = \frac{\sum_{i=1}^n(y_i-\hat{y_i})^2}{n} = \frac{\sum_{i=1}^n{e_i^2}}{n}$$
and the mean absolute error (MAE):

$$\mathrm{MAE} = \frac{\sum_{i=1}^n\left| y_i-\hat{y_i}\right|}{n} = \frac{\sum_{i=1}^n\left| e_i\right|}{n}$$
where $\hat{y_i}$ is the predicted value, given in our last single-unit layer, and $y_i$ is the actual value, the label.

# Part 3: k-fold cross validation

```{r setkFold, echo = TRUE, results = 'hide'}
k <- 4 # four groups
indices <- sample(1:nrow(train_data)) # randomize the training set before splitting for k-fold cross validation:
folds <- cut(indices, breaks = k, labels = FALSE) # divide the ordered indices into k intervals, labelled 1:k.
```

```{r kfold100, cache = T}
num_epochs <- 100
all_scores <- c() # An empty vector to store the results from evaluation

for (i in 1:k) {
  cat("processing fold #", i, "\n")
  # Prepare the validation data: data from partition # k
  val_indices <- which(folds == i, arr.ind = TRUE) 
  
  # validation set: the ith partition
  val_data <- train_data[val_indices,]
  val_targets <- train_targets[val_indices]
  
  # Training set: all other partitions
  partial_train_data <- train_data[-val_indices,]
  partial_train_targets <- train_targets[-val_indices]
  
  # Call our model function (see above)
  network <- build_model()
  
  # summary(model)
  # Train the model (in silent mode, verbose=0)
  network %>% fit(partial_train_data,
                  partial_train_targets,
                  epochs = num_epochs,
                  batch_size = 1,
                  verbose = 0)
                
  # Evaluate the model on the validation data
  results <- network %>% evaluate(val_data, val_targets, verbose = 0)
  all_scores <- c(all_scores, results$mean_absolute_error)
}  
```

We get 4 mae values

```{r allscores}
all_scores
```

### Training for 500 epochs

Let's try training the network for a bit longer: 500 epochs. To keep a record of how well the model did at each epoch, we will modify our training loop to save the per-epoch validation score log:

```{r clearMem}
# Some memory clean-up
K <- backend()
K$clear_session()
```

Train our the models:

```{r kfold500, echo = T, results = 'hide', cache = T}
num_epochs <- 500
all_mae_histories <- NULL # an empty object to cumulatively store the model metrics

for (i in 1:k) {
  cat("processing fold #", i, "\n")
  
  # Prepare the validation data: data from partition # k
  val_indices <- which(folds == i, arr.ind = TRUE)
  val_data <- train_data[val_indices,]
  val_targets <- train_targets[val_indices]
  
  # Prepare the training data: data from all other partitions
  partial_train_data <- train_data[-val_indices,]
  partial_train_targets <- train_targets[-val_indices]
  
  # Build the Keras model (already compiled)
  model <- build_model()
  
  # Train the model (in silent mode, verbose=0)
  history <- model %>% fit(partial_train_data, 
                           partial_train_targets,
                           validation_data = list(val_data, val_targets),
                           epochs = num_epochs, 
                           batch_size = 1, 
                           verbose = 0
  )
  mae_history <- history$metrics$val_mean_absolute_error
  all_mae_histories <- rbind(all_mae_histories, mae_history)
}
```

Calculate the average per-epoch MAE score for all folds:

```{r plot1}
average_mae_history <- data.frame(
  epoch = seq(1:ncol(all_mae_histories)),
  validation_mae = apply(all_mae_histories, 2, mean)
)

p <- ggplot(average_mae_history, aes(x = epoch, y = validation_mae))

p + 
  geom_point()

p + 
  geom_smooth(method = 'loess', se = FALSE)
```

According to this plot, it seems that validation MAE stops improving significantly after circa 80 epochs. Past that point, we start overfitting.

Once we are done tuning other parameters of our model (besides the number of epochs, we could also adjust the size of the hidden layers), we can train a final "production" model on all of the training data, with the best parameters, then look at its performance on the test data:

```{r runZ, echo = F, results = 'hide', cache = T}
# Get a fresh, compiled model.
model <- build_model()

# Train it on the entirety of the data.
model %>% fit(train_data, 
              train_targets,
              epochs = 80, 
              batch_size = 16, 
              verbose = 0)

result <- model %>% evaluate(test_data, test_targets)
```

```{r resultsZ}
result
```

We are still off by about `r round(result$mean_absolute_error * 1000)`.

## Alternatives: No Normalization

Let's imagine that we didn't normalize the input variables

```{r setupNone, echo = F, cache = T}
# Obtain the raw data
c(c(train_data, train_targets), c(test_data, test_targets)) %<-% dataset_boston_housing()

# Some memory clean-up
K <- backend()
K$clear_session()

num_epochs <- 500
all_mae_histories <- NULL # an empty object to cumulatively store the model metrics

for (i in 1:k) {
  cat("processing fold #", i, "\n")
  
  # Prepare the validation data: data from partition # k
  val_indices <- which(folds == i, arr.ind = TRUE)
  val_data <- train_data[val_indices,]
  val_targets <- train_targets[val_indices]
  
  # Prepare the training data: data from all other partitions
  partial_train_data <- train_data[-val_indices,]
  partial_train_targets <- train_targets[-val_indices]
  
  # Build the Keras model (already compiled)
  model <- build_model()
  
  # Train the model (in silent mode, verbose=0)
  history <- model %>% fit(partial_train_data, 
                           partial_train_targets,
                           validation_data = list(val_data, val_targets),
                           epochs = num_epochs, 
                           batch_size = 1, 
                           verbose = 0
  )
  mae_history <- history$metrics$val_mean_absolute_error
  all_mae_histories <- rbind(all_mae_histories, mae_history)
}
```

Calculate the average per-epoch MAE score for all folds:

```{r plot2, echo = F, cache = T}
average_mae_history <- data.frame(
  epoch = seq(1:ncol(all_mae_histories)),
  validation_mae = apply(all_mae_histories, 2, mean)
)

p <- ggplot(average_mae_history, aes(x = epoch, y = validation_mae))

p + 
  geom_point()

p + 
  geom_smooth(method = 'loess', se = FALSE)
```

The validation MAE stops improving significantly after circa 140 epochs.

```{r runNone, echo = F, results = 'hide', cache = T}
# Get a fresh, compiled model.
model <- build_model()

# Train it on the entirety of the data.
model %>% fit(train_data, 
              train_targets,
              epochs = 140, 
              batch_size = 16, 
              verbose = 0)

result_none <- model %>% evaluate(test_data, test_targets)
```

```{r resultsNone}
result_none
```

Now, without any normalization, we're off by about `r round(result_none$mean_absolute_error * 1000)`, compared to `r round(result$mean_absolute_error * 1000)` previously.

## Alternatives: 0-1 normalization 

How about if we did 0-1 normalization?

$$z_i=\frac{x_i-\min(x)}{\max(x)-\min(x)}$$

```{r setup01, echo = F, cache = T}
# Obtain the raw data
c(c(train_data, train_targets), c(test_data, test_targets)) %<-% dataset_boston_housing()

train_data <- apply(train_data, 2, function(x) (x-min(x))/(max(x)-min(x)))
test_data <- apply(test_data, 2, function(x) (x-min(x))/(max(x)-min(x)))

# Some memory clean-up
K <- backend()
K$clear_session()

num_epochs <- 500
all_mae_histories <- NULL # an empty object to cumulatively store the model metrics

for (i in 1:k) {
  cat("processing fold #", i, "\n")
  
  # Prepare the validation data: data from partition # k
  val_indices <- which(folds == i, arr.ind = TRUE)
  val_data <- train_data[val_indices,]
  val_targets <- train_targets[val_indices]
  
  # Prepare the training data: data from all other partitions
  partial_train_data <- train_data[-val_indices,]
  partial_train_targets <- train_targets[-val_indices]
  
  # Build the Keras model (already compiled)
  model <- build_model()
  
  # Train the model (in silent mode, verbose=0)
  history <- model %>% fit(partial_train_data, 
                           partial_train_targets,
                           validation_data = list(val_data, val_targets),
                           epochs = num_epochs, 
                           batch_size = 1, 
                           verbose = 0
  )
  mae_history <- history$metrics$val_mean_absolute_error
  all_mae_histories <- rbind(all_mae_histories, mae_history)
}
```

Calculate the average per-epoch MAE score for all folds:

```{r plot3, echo = F, cache = T}
average_mae_history <- data.frame(
  epoch = seq(1:ncol(all_mae_histories)),
  validation_mae = apply(all_mae_histories, 2, mean)
)

p <- ggplot(average_mae_history, aes(x = epoch, y = validation_mae))

p + 
  geom_point()

p + 
  geom_smooth(method = 'loess', se = FALSE)
```

The validation MAE stops improving significantly after circa 140 epochs.

```{r run01, echo = F, results = 'hide', cache = T}
# Get a fresh, compiled model.
model <- build_model()

# Train it on the entirety of the data.
model %>% fit(train_data, 
              train_targets,
              epochs = 140, 
              batch_size = 16, 
              verbose = 0)

result_01 <- model %>% evaluate(test_data, test_targets)
```

```{r results01}
result_01
```

Here, we're off by about `r round(result_01$mean_absolute_error * 1000)`, compared to `r round(result$mean_absolute_error * 1000)` with z scores.


































## Abalone Dataset

### Part 1: Data Preparation

| Variable       | Type       | Unit  | Description                 |
|----------------|------------|-------|-----------------------------|
| sex            | nominal    | --    | M, F, and I (infant)        |
| length         | continuous | mm    | Longest shell measurement   |
| diameter       | continuous | mm    | perpendicular to length     | 
| height         | continuous | mm    | with meat in shell          |
| whole_weight   | continuous | grams | whole abalone               |
| shucked_weight | continuous | grams | weight of meat              |
| viscera_weight | continuous | grams | gut weight (after bleeding) | 
| shell_weight   | continuous | grams | after being dried           | 
| rings          | integer    | --    | +1.5 gives the age in years |

The number of rings, variable `rings`, is the value to predict as either a continuous value or as a classification problem. 

this data set is also available in the `AppliedPredictiveModeling` package.

```{r eval = FALSE}
# load the library
library(AppliedPredictiveModeling)
data(abalone)
newdata <- abalone
dim(abalone)
head(abalone)

```


### Outline of terms

We'll perform a regression to predict a continouous response variable from 8 predictor variables. To accommodate for this different analytical problem, we'll use:

- Normalization for the input data: z scores
- Loss function: `mse`
- Metric: `mae`
- No final activation function (i.e. scalar)

And since we have a really small data set we'll have:

- A very simple network architecture, and
- K-fold crossvalidation.

### Obtain data &  Prepare data:

```{r eval = TRUE}

abalone_names <- c("Type",
                   "LongestShell",
                   "Diameter",
                   "Height",
                   "WholeWeight",
                   "WhuckedWeight",
                   "VisceraWeight",
                   "ShellWeight",
                   "Rings")

abalone <- read.csv("Abalone/abalone.data",
                    header = F,
                    col.names = abalone_names)



# Convert sex to integer :
abalone %>% 
  mutate(Type = as.integer(Type)) -> abalone

```

```{r}
glimpse(abalone)
```

```{r}
abalone %>% 
  group_by(Rings) %>% 
  summarise(n = n()) %>% 
  knitr::kable()
```

All values from 1-27 & 29 are present. The training and test set should contain at least one representative of each group.

### Examine data:

```{r}
tabplot::tableplot(abalone)
```

### plot the data anew:

```{r}
abalone %>% 
  select(-Rings) %>% 
  gather() %>%
  ggplot(aes(key, value)) +
  geom_jitter(shape = 1, alpha = 0.2)
```


```{r}
ggplot(abalone, aes(Rings)) +
  geom_bar() +
  scale_x_continuous("Number of Rings", breaks = 1:29) +
  coord_cartesian(expand = 0) +
  theme_minimal()

```

## Training and Test sets

```{r}
train_n <- round(0.8*nrow(abalone))
test_n <- round(0.2*nrow(abalone))
```

number of training instances n = `r train_n`.

number of test instances n = `r test_n`.

number of features d = `r ncol(abalone) - 1`.

number of classes K = `r length(unique(abalone$rings))`.

### Split up training and test


```{r}

# Convert to a matrix:
abalone <- as.matrix(abalone)

# add additional vector to make match even
add_on_matrix <- matrix(999, ncol = 8, nrow = 28)
add_on_vector <- c(1:27,29)

set.seed(136)
train_index <- sample(seq_len(nrow(abalone)), train_n)

train_data <- unname(abalone[train_index, -9])
train_data <- rbind(train_data, add_on_matrix)

train_labels <- unname(abalone[train_index, 9])
train_labels <- c(train_labels, add_on_vector)

test_data <- unname(abalone[-train_index, -9])
test_data <- rbind(test_data, add_on_matrix)

test_labels <- unname(abalone[-train_index, 9])
test_labels <- c(test_labels, add_on_vector)

rm(abalone, abalone_names, train_n, test_n, train_index)

```


```{r}
str(train_data)
str(test_data)

```

## Labels

The `_labels` objects contain the news wire labels. Each newswire can only have one *label* (i.e. "sigle-label"), from a total of 46 possible *classes* (i.e. "multi-class"). The classes are just given numerical values (0 - 45), it doesn't matter what they are actually called, although that information would be helpful in understanding mis-labeling.

```{r}
table(train_labels)
```

```{r}
table(test_labels)
```

Some classes are very common, which we'll see play out in our confusion matrix below 

```{r plotLabelsPre}
# Note plyr not dplyr here. I'm just using a shortcut
library(ggplot2)
train_labels %>% 
  plyr::count() %>%
  ggplot(aes(x, freq)) +
  geom_col()
```

The distribution of the test and training set should be roughly equivalent, so let's have a look. 

```{r}
data.frame(x = train_labels) %>% 
  group_by(x) %>% 
  summarise(train_freq = 100*n()/length(train_labels)) -> train_labels_df

data.frame(x  = test_labels) %>% 
  group_by(x) %>% 
  summarise(test_freq = 100 * n()/length(test_labels)) %>% 
  inner_join(train_labels_df, by="x") %>% 
  gather(key, value, -x) %>% 
  ggplot(aes(x, value, fill = key)) +
  geom_col(position = "dodge") +
  # scale_y_continuous("Percentage", limits = c(0,20), expand = c(0,0)) +
  # scale_x_continuous("Label", breaks = 0:45, expand = c(0,0)) +
  scale_fill_manual("", labels = c("test","train"), values = c("#AEA5D0", "#54C8B7")) +
  theme_classic() +
  theme(legend.position = c(0.8, 0.8),
        axis.line.x = element_blank(),
        axis.text = element_text(colour = "black"))
```

We treat these just like how we treated the MNIST labels in the previous unit. We make the format match the output we expect to get from softmax so that we can make a direct comparison.

```{r prepLabels}
train_labels_vec <- to_categorical(train_labels)
test_labels_vec <- to_categorical(test_labels)
```

```{r}
colSums(test_labels_vec)
colSums(train_labels_vec)
```


```{r strLabelsPost}
str(train_labels_vec)
str(test_labels_vec)
```

## As a Classification Problem

### Part 2: Define Network

```{r architecture}
network <- keras_model_sequential() %>% 
  layer_dense(units = 2^5, activation = "relu", input_shape = 8) %>% 
  # layer_dense(units = 2^5, activation = "relu") %>% 
  layer_dense(units = 30, activation = "softmax")

```

### View a summary of the network

```{r summary}
summary(network)
```

### Compile

```{r compile}
network %>% compile(
  optimizer = "rmsprop",
  loss = "categorical_crossentropy",
  metrics = c("accuracy")
)
```

## Part 3: Validate our approach

Let's set apart 20% of the samples in our training data to use as a validation set:

```{r}

index <- 1:(0.2*nrow(train_data))

val_data_vec <- train_data[index,]
train_data_vec <- train_data[-index,]

train_labels_vec_original <- train_labels_vec
val_labels_vec <- train_labels_vec[index,]
train_labels_vec = train_labels_vec[-index,]

```

Now let's train our network for 20 epochs:

```{r echo=TRUE, results = "hide", warning = FALSE}
history <- network %>% fit(
  train_data_vec,
  train_labels_vec,
  epochs = 20,
  batch_size = 512,
  validation_data = list(val_data_vec, val_labels_vec)
)
```

Let's display its loss and accuracy curves:

```{r}
plot(history)
```

The network begins to overfit after nine epochs. Let's train a new network from scratch for nine epochs and then evaluate it on the test set.

```{r, echo=TRUE, results='hide'}
network <- keras_model_sequential() %>% 
  layer_dense(units = 2^5, activation = "relu", input_shape = 8) %>% 
  # layer_dense(units = 2^5, activation = "relu") %>% 
  layer_dense(units = 30, activation = "softmax")
  
network %>% compile(
  optimizer = "rmsprop",
  loss = "categorical_crossentropy",
  metrics = c("accuracy")
)

history <- network %>% fit(
  train_data_vec,
  train_labels_vec,
  epochs = 9,
  batch_size = 512,
  validation_data = list(val_data_vec, val_labels_vec)
)
```

# Regression using sparse categorical crossentropy

Alternatively, we could have just used the original integer values. To showcase this, let's create a new network, `network_int`, so that we don't mix up our results. The network architecture is the same:

```{r}
network_int <- keras_model_sequential() %>% 
  layer_dense(units = 2^5, activation = "relu", input_shape = 8) %>% 
  # layer_dense(units = 2^5, activation = "relu") %>% 
  layer_dense(units = 30, activation = "softmax")
```

Here, the only thing we need to chance is the loss function. `categorical_crossentropy`, expects the labels to follow a categorical encoding, but `sparse_categorical_crossentropy` expects integer labels. 

```{r}
network_int %>% compile(
  optimizer = "rmsprop",
  loss = "sparse_categorical_crossentropy",
  metrics = c("accuracy")
)
```

Before we train the model, let's make a validation set, like we did above. We'll use the original training set for this.

```{r}
val_train_labels <- train_labels[index]
train_labels <- train_labels[-index]
```

Now let's train our model `network_int` using the integer data, instead of the vectorized data:

```{r}
history_int <- network_int %>% fit(
  train_data_vec,
  train_labels,
  epochs = 9,
  batch_size = 512,
  validation_split = list(val_data_vec, val_train_labels)
)
```

This new loss function is mathematically the same as `categorical_crossentropy`. It just has a different interface. When we look at our metrics below we'll use the original model, that accessed the vectorized data. If you want to use `network_int` make sure you use the original integer labels of the test set, `test_labels`, not `test_labels_vec`. 

## Part 5: Check output

Let's return to our original model using the vectorized data:

