---
title: "The Boston House Price Dataset"
subtitle: "Classic Machine Learning Approaches"
author: "Rick Scavetta"
output:
  html_document:
  fig_caption: true
  toc: true
  toc_float:
    collapsed: false
    smooth_scroll: false
  toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, eval = TRUE)

# Initialize packages
library(keras)
library(tidyverse)
library(caret)
library(randomForest)
library(kableExtra)
```

# {.tabset .tabset-fade .tabset-pills}

## Introduction

Although this workshop focuses on deep learning, I want to put it into context with some methods that you should already be familiar with.

In this script, we'll also get familiar with the Boston Housing Price dataset.

## Load data

For this example, I'm going to use the exact same data that we'll see later in deep learning so we can directly compare the results.

```{r}
# Read the dataset
boston <- read.csv("../data/boston_keras.csv")
```

We have 14 variables. The first 13 are the predictor variables and that last, `MEDV` is the response

| Variable  | Description                                                               |
|:----------|:--------------------------------------------------------------------------|
| `CRIM`    | Per capita crime rate by town                                             |
| `ZN`      | Proportion of residential land zoned for lots over 25,000 sq.ft           |
| `INDUS`   | Proportion of non-retail business acres per town                          |
| `CHAS`    | Charles River dummy variable (1 if tract bounds river; else 0)            |
| `NOX`     | Nitric oxides concentration (parts per 10 million)                        |
| `RM`      | Average number of rooms per dwelling                                      |
| `AGE`     | Proportion of owner-occupied units built prior to 1940                    |
| `DIS`     | Weighted distances to five Boston employment centres                      |
| `RAD`     | Index of accessibility to radial highways                                 |
| `TAX`     | Full-value property-tax rate per $10,000                                  |
| `PTRATIO` | Pupil-teacher ratio by town                                               |
| `B`       | $1000 * (Bk - 0.63)^2$ where $Bk$ is the proportion of blacks by town |
| `LSTAT`   | % lower status of the population                                          |
| `MEDV`    | Median value of owner-occupied homes in $1000’s                           |

To give you an idea of the range of the data:

```{r}
summary(boston)
```

## ML preparation

### Check Correlation

Each predictor variable against the response variable:

```{r echo = FALSE}

data.frame(r = cor(boston[,-14],boston$MEDV)) %>% 
  rownames_to_column() %>% 
  rename(variable = rowname) %>% 
  arrange(r) %>% 
  mutate(variable = as_factor(variable)) %>% 
  ggplot(aes(r, variable)) +
  geom_vline(xintercept = 0, col = "dark red", linetype = 2) +
  geom_point() +
  scale_x_continuous("r", limits = c(-1,1), expand = c(0,0))
  
```

In practice we may decide, using criteria such as r, that some variables are more informative. In this case we'll just take all the variables.

### Check Variance

Assess presence of zero of near-zero variance variables:

```{r}
nzv <- nearZeroVar(boston, saveMetrics = TRUE)
```

There are `r sum(nzv$nzv)` zero variance of near-zero variance variables.

### Create test set

This split reflects what we'll see in `keras`.

```{r}
# as per the keras data set:
index <- 1:404

training <- boston[index,]
testing <- boston[-index,]
```

### Z-score Scaling

We'll perform a Z-score transformation on each predictor variable, after splitting, as we will do later in deep learning. All variables will have a mean of 0.

```{r}
# Using dplyr to keep data frame structure:
training %>% 
  mutate_at(vars(-MEDV), scale) -> training

testing %>% 
  mutate_at(vars(-MEDV), scale) -> testing
```

## GLM

Linear models take the form:

$$Y_i = \beta_0 + \beta_1 X_{i1} + \beta_2 X_{i2} + \ldots + \beta_p X_{ip} + \epsilon_i$$. The coefficients are:

Since we are assuming normal distributions, `glm()` and `lm()` perform the same. The coefficients:

```{r}
fit_lm <- lm(MEDV~.,data = training)
```


```{r echo = FALSE}
data.frame(coef = round(fit_lm$coefficients,2)) %>% 
  rownames_to_column() %>% 
  rename(variable = rowname) %>% 
  filter(variable != "(Intercept)") %>%  
  arrange(coef) %>% 
  mutate(variable = as_factor(variable)) %>% 
  ggplot(aes(coef, variable)) +
  geom_vline(xintercept = 0, col = "dark red", linetype = 2) +
  geom_point() +
  scale_x_continuous("r", limits = c(-4.2,4.2), expand = c(0,0))
```


```{r}
#predict on test set
pred_lm <- predict(fit_lm, newdata = testing)

MAE_lm <- sum(abs(pred_lm - testing$MEDV))/102
```

Our measure for the error will be the mean absolute error (MAE):

$$\mathrm{MAE} = \frac{\sum_{i=1}^n\left| y_i-\hat{y_i}\right|}{n} = \frac{\sum_{i=1}^n\left| e_i\right|}{n}$$

where $\hat{y_i}$ is the predicted value and $y_i$ is the actual value, the label. We'll see this again in deep learning and it's a more intuitive unit than the root-mean-square error (RMSE), which is also common. Recall that the RMSE is just the square of the the MSE:

$$\operatorname{MSE} = \frac{\sum_{i=1}^n(y_i-\hat{y_i})^2}{n} = \frac{\sum_{i=1}^n{e_i^2}}{n}$$

$$\operatorname{RMSE} = \sqrt{MSE}$$

the MSE will also make a reappearance in deep learning as our loss function.

The MAE, using the linear model, is `r MAE_lm`. In dollar amounts, we are off by $`r round(MAE_lm * 1000, 2)`.

## Random Forest

Let's give it another go using a random forest.

```{r}
fit_rf <- randomForest(MEDV ~ ., data = training)

pred_rf <- predict(fit_rf, testing)

MAE_rf <- sum(abs(pred_rf - testing$MEDV))/102

```

In this case the MAE is `r MAE_rf`. In dollar amounts, we are off by $`r round(MAE_rf * 1000)`, much better than before.

## Visualizing output

Let's put this into context. The predictions for each model are compared to the actual data. The diagonal line is 1:1 equivalency.

```{r echo = FALSE}
data.frame(Actual = testing$MEDV,
           `GLM` = pred_lm,
           `Random Forest` = pred_rf) %>% 
  gather(Measure, Prediction, -Actual) %>% 
  ggplot(aes(Actual, Prediction)) +
  geom_point(shape = 16, alpha = 0.65) +
  geom_abline(slope = 1, intercept = 0, col = "dark red") +
  coord_fixed() +
  facet_grid(. ~ Measure) +
  theme_classic() +
  theme(axis.text = element_text(colour = "black"),
        strip.background = element_rect(colour = NA, fill = "gray92"))
```

```{r, eval = FALSE, echo = FALSE}
data.frame(Actual = testing$MEDV,
           `GLM` = pred_lm,
           `Random Forest` = pred_rf) %>% 
  rio::export("ClassicML.csv")
```

The results are saved in the `ClassicML.csv` file so that we can compare them with the results from deep learning.