---
title: "Introduction to Deep Learning"
subtitle: "Classification -- Binary"
author: "Rick Scavetta"
output:
  html_document:
  fig_caption: true
  toc: true
  toc_float:
    collapsed: false
    smooth_scroll: false
  toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, eval = TRUE)

# Initialize packages
library(keras)
library(tidyverse)
```

# {.tabset .tabset-fade .tabset-pills}

## Functions & Take-home Message

For binary classification, we try to predict one of two possible classes for the response variable.

### Take-home Message

In this case study, we'll perform a binary classification to predict one of two possible outcomes from 13 predictor variables. 

All the actual function will remain the same as before, except:

| Name                | Function              |
|:--------------------|:----------------------|
| `predict_classes()` | Predict classes using a model and test data. |

We'll use the following functions in our network arguments:

| Name                           | Function              |
|:-------------------------------|:----------------------|
| Loss function                  | `binary_crossentropy` |
| Metric                         | `accuracy`            |
| last-layer activation function | `sigmoid`             |

## Prepare the training data:

Obtain data

```{r}
source("Boston_Z.R")
```


The target, response variable is continuous. For this example, we'll pretend that it's binary.

```{r strTargets}
str(train_targets)
str(test_targets)
```

```{r}

train_targets <- as.numeric(cut(train_targets, seq(0,50,25))) - 1
test_targets <- as.numeric(cut(test_targets, seq(0,50,25))) - 1

table(train_targets)
table(test_targets)

```

## Part 2: Define Network

### Define the network

Here we specify the final activation function. We're going to use the sigmoid activation function, which will return a single value. That matches the format of our labels.

```{r architecture}
network <- _______ %>% 
  _______ %>% 
  _______ %>% 
  _______
```

### View a summary of the network

```{r summary}
summary(network)
```

### Compile

Instead of `categorical_crossentropy` we're going to use `binary_crossentropy` since we only have two possible classes.

```{r compile}
network %>% compile(
  optimizer = "rmsprop",
  loss = _______,
  metrics = _______
)
```

## Part 3: Train Network

Now let's train our network for 100 epochs:

```{r echo=TRUE, results = "hide", warning = FALSE}
history <- network %>% 
  _______(train_data,
      train_targets,
      epochs = 100,
      batch_size = 16
  )
```

Let's display its loss and accuracy curves:

```{r}
plot(history)
```

The network begins to over-fit after about 30 epochs. Let's train a new network from scratch for four epochs and then evaluate it on the test set.

```{r, echo=TRUE, results='hide'}
network <- keras_model_sequential() %>% 
  layer_dense(units = 2^6, activation = "relu", input_shape = 13) %>% 
  layer_dense(units = 2^6, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")
  
network %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history <- network %>% fit(
  train_data,
  train_targets,
  epochs = _______,
  batch_size = 16
)
```

## Part 4: Check output

### Metrics

```{r metrics}
metrics <- network %>% _______(test_data, test_targets)
```

```{r}
metrics
```

The accuracy is `r metrics$acc` and the error rate, i.e. the incorrect calling, is `r 1 - metrics$acc`.

### Predictions

```{r predictions}
network %>% _______(test_data[1:10,])
```

```{r allPredictions}
predictions <- network %>% predict_classes(test_data)
actual <- unlist(test_targets)
totalmisses <- sum(predictions != actual)
totalmisses
```

A total of `r totalmisses` mismatches occurred.

### Confusion Matrix

```{r confusion, echo = F}
data.frame(target = actual,
           prediction = predictions) %>% 
  filter(target != prediction) %>% 
  group_by(target, prediction) %>%
  count() %>%
  filter(n > 1) %>% 
  ggplot(aes(target, prediction, size = n)) +
  geom_point(shape = 15, col = "#9F92C6") +
  scale_x_continuous("Actual Target", breaks = 0:1, labels = c("low", "high")) +
  scale_y_continuous("Prediction", breaks = 0:1, labels = c("low", "high")) +
  scale_size_area(max_size = 10) +
  coord_fixed() +
  ggtitle(paste(totalmisses, "mismatches")) +
  theme_classic() +
  theme(rect = element_blank(),
        axis.line = element_blank(),
        axis.text = element_text(colour = "black"))
```
