---
title: "Session 1 Exercise: Abaolone as Regression"
subtitle: "Single-label, Multi-class Classification"
author: "Rick Scavetta"
output:
  html_document:
  fig_caption: true
  toc: true
  toc_float:
    collapsed: false
    smooth_scroll: false
  toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Initialize package
library(keras)
library(tidyverse)
```

# {.tabset .tabset-fade .tabset-pills}

## Obtain data

```{r data, warning = FALSE}

data_dir <- "data/"

train_data <- read_csv(paste0(data_dir, "abalone_data.csv"))
train_labels <- read_csv(paste0(data_dir, "abalone_labels_4cat.csv"))
train_labels <- train_labels$Rings

set.seed(136)
index <- sample(1:nrow(train_data), 0.2*nrow(train_data))

test_data <- train_data[index,]
test_labels <- train_labels[index]

train_data <- train_data[-index,]
train_labels <- train_labels[-index]

```

## Prepare

Examine data:

Our predictor variables:

```{r strDataPre}
str(train_data)
str(test_data)
```

### Prepare the data:

The target, response variable:

```{r strTargets}
str(train_labels)
```

Convert z-scores:

```{r zScores, cache = T}
# parameters for Scaling:
mean <- colMeans(train_data) # mean of each column
std <- apply(train_data, 2, sd) # stdev of each column

# Calculate feature-wise (within-variable) z-scores: (x - mean)/std
train_data <- scale(train_data, center = mean, scale = std)
test_data <- scale(test_data, center = mean, scale = std)
```

### Prepare labels:

```{r strLabelsPre}
str(train_labels)
sort(unique(train_labels))
```

Some classes are very common, which we'll see play out in our confusion matrix below 

```{r plotLabelsPre}
# Note plyr not dplyr here. I'm just using a shortcut
library(ggplot2)
train_labels %>% 
  plyr::count() %>%
  ggplot(aes(x, freq)) +
  geom_col()
```

The distribution of the test and training set should be roughly equivalent, so let's have a look. 

```{r}
data.frame(x = train_labels) %>% 
  group_by(x) %>% 
  summarise(train_freq = 100 * n()/length(train_data)) -> train_labels_df

data.frame(x  = test_labels) %>% 
  group_by(x) %>% 
  summarise(train_freq = 100 * n()/length(test_data)) %>% 
  inner_join(train_labels_df, by="x") %>% 
  gather(key, value, -x) %>% 
  ggplot(aes(x, value, fill = key)) +
  geom_col(position = "dodge") +
  scale_y_continuous("Percentage", limits = c(0,40), expand = c(0,0)) +
  scale_x_continuous("Label", breaks = 0:45, expand = c(0,0)) +
  scale_fill_manual("", labels = c("test","train"), values = c("#AEA5D0", "#54C8B7")) +
  theme_classic() +
  theme(legend.position = c(0.8, 0.8),
        axis.line.x = element_blank(),
        axis.text = element_text(colour = "black"))
```

## Part 2: Define Network

### Define the network

```{r architecture}
network <- ______ %>% 
  ______ %>% 
  ______ %>% 
  ______

```

```{r summary}
summary(network)
```

### Compile

```{r compile}
network %>% ______(
  optimizer = "rmsprop",
  ______ = ______,
  ______ = ______
)
```

## Train

```{r echo=TRUE, results = "hide", warning = FALSE}
history <- network %>% ______(
  ______,
  ______,
  ______ = ______,
  ______ = ______
)
```

Let's display its loss and accuracy curves:

```{r}
plot(history)
```


## Part 4: Check output

Let's return to our original model using the vectorized data:

$## Metrics

```{r metrics}
metrics <- network %>% ______(______, ______)
```

```{r}
metrics
```

##$ Predictions

```{r predictions}
network %>% ______(______)
```

```{r allPredictions}
predictions <- network %>% ______(______)
actual <- unlist(test_labels)
totalmisses <- sum(predictions != actual)
```

### Confusion Matrix

```{r confusion, echo = F}
data.frame(target = actual,
           prediction = predictions) %>% 
  filter(target != prediction) %>% 
  group_by(target, prediction) %>%
  count() %>%
  ungroup() %>%
  mutate(perc = n/nrow(.)*100) %>% 
  filter(n > 1) %>% 
  ggplot(aes(target, prediction, size = n)) +
  geom_point(shape = 15, col = "#9F92C6") +
  scale_x_continuous("Actual Target", breaks = 1:4, limits = c(1,4)) +
  scale_y_continuous("Prediction", breaks = 1:4, limits = c(1,4)) +
  scale_size_area(breaks = c(2,5,10,15), max_size = 5) +
  coord_fixed() +
  ggtitle(paste(totalmisses, "mismatches")) +
  theme_classic() +
  theme(rect = element_blank(),
        axis.line = element_blank(),
        axis.text = element_text(colour = "black"))

```
