---
title: "Reuters -- Single-label, Multi-class Classification with Text"
subtitle: "Scenario 1: One-hot Encoding"
author: "Rick Scavetta"
output:
  html_document:
  fig_caption: true
  toc: true
  toc_float:
    collapsed: false
    smooth_scroll: false
  toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

# Initialize package
library(keras)
library(tidyverse)
```

# {.tabset .tabset-fade .tabset-pills}

## Learning Goals & Functions

### Learning Goals

Text analysis using one-hot encoding

### Functions in this session:

Here we'll see some basic ways of working with text, using one-hot encoding.

| Function                       | Description                                                                         |
|:-------------------------------|:------------------------------------------------------------------------------------|
| `dataset_reuters()`            | Import the Reuters Newswires dataset.                                               |
| `dataset_reuters_word_index()` | Reuters dataset word index.                                                         |
| `vectorize_sequences()`        | Custom function to produce one-hot encoded data.                                    |
| `to_categorical()`             | Vectorize labels for use with softmax and `categorical_crossentropy` loss function. |

## Part 1: Data Preparation

For multi-class classification, we previously use `sparse_cateogircal_crossentropy`, here we'll see  `categorical_crossentropy`.

### Obtain data

```{r data, warning = FALSE}
c(c(train_data, train_labels), c(test_data, test_labels)) %<-% _______(num_words = 10000)
```

### Examine data:

```{r strImagesPre}
length(train_data)
length(test_data)
```

An example of the data

```{r}
train_data[[1]]
```

Map values back onto the lexicon which is stored as a named list. Each item in the list is an integer vector of length one. This number corresponds to the position in the word count list and the name of the vector is the actual word. 

```{r}
_______() %>% 
  unlist() %>%                      # produce a vector
  sort() %>%                        # put them in order 
  names() -> word_index             # take the ordered names

# The indices are offset by 3 since 0, 1, and 2 are reserved 
# for "padding", "start of sequence", and "unknown"
library(purrr)
train_data[[1]] %>% 
  map(~ ifelse(.x >= 3, word_index[.x - 3], "?")) %>% 
  as_vector() %>% 
  cat()
```

### Prepare the data:

We'll use one-hot encoding:

```{r normImages}
vectorize_sequences <- function(sequences, dimension = 10000) {
  # Create a matrix of 0s
  results <- matrix(0, nrow = length(sequences), ncol = dimension)

  # Populate the matrix with 1s
  for (i in 1:length(sequences))
    results[i, sequences[[i]]] <- 1
  results
}

train_data_vec <- _______(train_data)
test_data_vec <- _______(test_data)
```

Let's look at the first example from the training set. Recall that these are the index positions of the words 

```{r}
train_example <- sort(unique(train_data[[1]]))
train_example
```

Now we have a large matrix, where each row is 10000 elements long. Wherever we have a value in the above data set, the matrix has a 1

```{r}
# Just the first 100 values in the first entry (row)
train_data_vec[1,1:100]
```

In one-hot encoding, we only get 1s and 0s. The position of the 1s corresponds to the indices above:

```{r}
which(as.logical(train_data_vec[1,]))
```

### Prepare labels:

The `_labels` objects contain the news wire labels. Each newswires can only have one *label* (i.e. "single-label"), from a total of 46 possible *classes* (i.e. "multi-class"). The classes are just given numerical values (0 - 45), it doesn't matter what they are actually called, although that information would be helpful in understanding mis-labeling.

```{r strLabelsPre}
str(train_labels)
```

Some classes are very common, which we'll see play out in our confusion matrix below 

```{r plotLabelsPre}
# Note plyr not dplyr here. I'm just using a shortcut
library(ggplot2)
train_labels %>% 
  plyr::count() %>%
  ggplot(aes(x, freq)) +
  geom_col()
```

The distribution of the test and training set should be roughly equivalent, so let's have a look. 

```{r}
data.frame(x = train_labels) %>% 
  group_by(x) %>% 
  summarise(train_freq = 100 * n()/length(train_data)) -> train_labels_df

data.frame(x  = test_labels) %>% 
  group_by(x) %>% 
  summarise(train_freq = 100 * n()/length(test_data)) %>% 
  inner_join(train_labels_df, by="x") %>% 
  gather(key, value, -x) %>% 
  ggplot(aes(x, value, fill = key)) +
  geom_col(position = "dodge") +
  scale_y_continuous("Percentage", limits = c(0,40), expand = c(0,0)) +
  scale_x_continuous("Label", breaks = 0:45, expand = c(0,0)) +
  scale_fill_manual("", labels = c("test","train"), values = c("#AEA5D0", "#54C8B7")) +
  theme_classic() +
  theme(legend.position = c(0.8, 0.8),
        axis.line.x = element_blank(),
        axis.text = element_text(colour = "black"))
```

We make the format match the output we expect to get from softmax so that we can make a direct comparison.

```{r prepLabels}
train_labels_vec <- to_categorical(train_labels)
test_labels_vec <- to_categorical(test_labels)
```

```{r strLabelsPost}
str(train_labels_vec)
str(test_labels_vec)
```

Notice the similarity to how we prepared our training data with one-hot encoding, both are sparse matrices. Each row in the training data contains a 1 at the position where that word is present, and in the labels, each row contains a maximum of a single 1, indicating the class. 

## Part 2: Define Network

### Define the network

You shoud be familiar with this step by now.

```{r architecture}
network <- _______() %>% 
  _______(units = _______, activation = _______, input_shape = _______) %>% 
  _______(units = _______, activation = v) %>% 
  _______(units = _______, activation = _______)
```

### View a summary of the network

```{r summary}
summary(network)
```

### Compile

Which funcitons will we use here?

```{r compile}
network %>% compile(
  optimizer = "rmsprop",
  loss = _______,
  metrics = _______
)
```

## Part 3: Validate our approach

Let's set apart 1,000 samples in our training data to use as a validation set:

```{r}
index <- 1:1000
val_data_vec <- train_data_vec[index,]
train_data_vec <- train_data_vec[-index,]
val_labels_vec <- train_labels_vec[index,]
train_labels_vec = train_labels_vec[-index,]
```

Now let's train our network for 20 epochs:

Which objects to use for training and validation?

```{r echo=TRUE, results = "hide", warning = FALSE}
history <- network %>% fit(
  _______,
  _______,
  epochs = 20,
  batch_size = 512,
  validation_data = list(_______, v)
)
```

Let's display its loss and accuracy curves:

```{r}
plot(history)
```


```{r, echo=TRUE, results='hide'}
network <- keras_model_sequential() %>% 
  layer_dense(units = 64, activation = "relu", input_shape = c(10000)) %>% 
  layer_dense(units = 64, activation = "relu") %>% 
  layer_dense(units = 46, activation = "softmax")
  
network %>% compile(
  optimizer = "rmsprop",
  loss = "categorical_crossentropy",
  metrics = c("accuracy")
)
history <- network %>% fit(
  train_data_vec,
  train_labels_vec,
  epochs = _______,
  batch_size = 512,
  validation_data = list(val_data_vec, val_labels_vec)
)
```

In the first session, we saw multi-class classification where the values were encoded as integers. For that purpose we used `sparse_categorical_crossentropy` as the loss function. Here we use softmax as the final layer activation function because we have vectorized the labels.

## Part 4: Check output

Let's return to our original model using the vectorized data:

### Metrics

```{r metrics}

# Recall the original integer data:
test_labels[1]

# We vectorized it to get 46 values, all zero except 1 where our value is
# This matches the out put from the softmax function.
test_labels_vec[1,] 

metrics <- network %>% evaluate(test_data_vec, test_labels_vec)
```

```{r}
metrics
```

Our accuracy is `r metrics$acc` and the error rate, i.e. incorrect calling, is `r 1 - metrics$acc`.

### Predictions

Predict classes will give us the integer values from the test data. What is the `predict()` function to use?

```{r predictions}
predictions <- network %>% _______(test_data_vec)
predictions[1:10]
```

```{r allPredictions}
actual <- unlist(test_labels)
totalmisses <- sum(predictions != actual)
```

This means we have a total of `r totalmisses` misses.

### Confusion Matrix

```{r confusion, echo = F}

data.frame(target = actual,
           prediction = predictions) %>% 
  filter(target != prediction) %>% 
  group_by(target, prediction) %>%
  count() %>%
  ungroup() %>%
  mutate(perc = n/nrow(.)*100) %>% 
  filter(n > 1) %>% 
  ggplot(aes(target, prediction, size = n)) +
  geom_point(shape = 15, col = "#9F92C6") +
  scale_x_continuous("Actual Target", breaks = 0:45, limits = c(0,45)) +
  scale_y_continuous("Prediction", breaks = 0:45, limits = c(0,45)) +
  scale_size_area(breaks = c(2,5,10,15), max_size = 5) +
  coord_fixed() +
  ggtitle(paste(totalmisses, "mismatches")) +
  theme_classic() +
  theme(rect = element_blank(),
        axis.line = element_blank(),
        axis.text = element_text(colour = "black"))
```
