---
title: "Reuters -- Single-label, Multi-class Classification with Text"
subtitle: "Scenario 2: Word Embeddings"
author: "Rick Scavetta"
output:
  html_document:
  fig_caption: true
  toc: true
  toc_float:
    collapsed: false
    smooth_scroll: false
  toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

# Initialize package
# install.packages(keras)
library(keras)
library(tidyverse)
```

# {.tabset .tabset-fade .tabset-pills}

## Learning Goals & Functions

### Learning Goals

Text analysis using our own word embeddings.

### Functions in this session:

Here we'll see some basic ways of working with text, using one-hot encoding.

| Function              | Description                               |
|:----------------------|:------------------------------------------|
| `pad_sequences()`    | Make all instances the same length by padding them. |
| `layer_embedding()`    | Add an embedding layer to a model. |

## Obtain data

### Loading the Reuters data set for use with an embedding layer

Restrict the newswires to the top 10,000 most common words then cut off the reviews after only 20 words to make the situation a bit simpler. Load the data set and pad the sequences.

```{r}
# Number of most common words to consider as features
max_features <- 10000

# Loads the data as lists of integers
c(c(train_data, train_labels), c(test_data, test_labels)) %<-% _______(num_words = max_features)

# # Cut off the text after 20 words (i.e. among the max_features most common words)
maxlen <- 20

# Turns the lists of integers into a 2D integer tensor of shape (samples, maxlen)
train_data <- _______(train_data, maxlen = maxlen)
test_data <- _______(test_data, maxlen = maxlen)
```

```{r}
dim(test_data)
```

## Using an embedding layer

Embedding layer plus classifier on the Reuters dataset

The network will:

1. Learn 8-dimensional embeddings for each of the 10,000 words, 
2. Turn the input integer sequences (2D integer tensor) into embedded sequences (3D float tensor),
3. Flatten the tensor to 2D, and
4. Train a single dense layer on top for classification.

```{r}
# Part 1 & 2:
# Specify the maximum input length to the embedding layer so you can later flatten the embedded inputs.
network <- _______() %>%
  _______(input_dim = _______, 
                  output_dim = 8,
                  input_length = maxlen)

# Part 3:
# After the embedding layer, the activations have shape (samples, maxlen, 8).
# Flattens the 3D tensor of embeddings into a 2D tensor of shape (samples, maxlen * 8)
network <- network %>% 
  _______()

# Part 4:
# Adds the classifier on top
network <- network %>%
  _______(units = _______, activation = _______)

```

Compile the model

```{r}
network %>% compile(
  optimizer = "rmsprop",
  loss = _______,
  metrics = _______
)

summary(network)
```


Train the model. Take 20% for validation.

```{r}
history <- network %>% fit(
  train_data, 
  train_labels,
  epochs = 100,
  batch_size = 512,
  validation_split = _______
)
```


```{r}
plot(history)
```

Let's train until 30-40 epochs:

```{r}

network <- keras_model_sequential() %>%
  layer_embedding(input_dim = 10000, 
                  output_dim = 8,
                  input_length = maxlen) %>% 
  layer_flatten() %>%
  layer_dense(units = 46, activation = "softmax")

network %>% compile(
  optimizer = "rmsprop",
  loss = "sparse_categorical_crossentropy",
  metrics = "accuracy"
)

history <- network %>% fit(
  train_data, 
  train_labels,
  epochs = 35,
  batch_size = 512,
  validation_split = 0.2
)
```

```{r}
# Peak validation accuracy:
history$metrics$val_acc[35]
```

